---
title: "PULSE_EW-NT-20221019"
author: "Nate Thomas"
date: "2022-10-19"
output: html_document
---

```{r}
library("dplyr")
```

NIOSH WSH (PAR-20-312) Project - Household PULSE Data Analysis
Data: HPS data phase 3.1-3.5 (week 28-48)
Task#1: Essential work settings - counts 06/20/2022 ~
Task#2: Shot by EW 8/22/2022 ~

Table of Contents
1. Read data folders
2. Clean and prepare data
2.1 Merge data and wieghts
2.2 Clean data and create new variables - independent variables

# 1. Read Data

```{r}
#initialize empty data list
raw_data <- data.frame()
raw_weights <- data.frame()

#iterate through data folder to pull in each week
for (i in 1:(length(list.dirs('./Data'))-1)){
  dir = list.dirs('./Data')[i+1]
  if ((i+27)<41) {
    raw_data = dplyr::bind_rows(raw_data,read.csv(paste(dir,'/pulse2021_puf_',i+27,'.csv',sep='')))
    print(paste(dir,'/pulse2021_puf_',i+27,'.csv',sep=''))
    raw_weights = dplyr::bind_rows(raw_weights,read.csv(paste(dir,'/pulse2021_repwgt_puf_',i+27,'.csv',sep='')))
  }
  else {
    raw_data = dplyr::bind_rows(raw_data,read.csv(paste(dir,'/pulse2022_puf_',i+27,'.csv',sep='')))
    print(paste(dir,'/pulse2022_puf_',i+27,'.csv',sep=''))
    raw_weights = dplyr::bind_rows(raw_weights,read.csv(paste(dir,'/pulse2022_repwgt_puf_',i+27,'.csv',sep='')))
  }
}

#save.image()
```

# 2. Clean and prepare data

```{r}
#Year 2021 for weeks 28-40, otherwise 2022
df <- raw_data %>% 
  mutate(yr = ifelse(WEEK < 41,2021,2022))
```

## 2.1 Merge data and wieghts

```{r}
df_joined <- df %>%
  inner_join(raw_weights, by = c('SCRAM','WEEK'))
```

## 2.2 Keep only variables needed

```{r}
df_selected <- df_joined %>%
  mutate(`PWEIGHT1-PWEIGHT80` = PWEIGHT1-PWEIGHT80) %>%
  select(## UNIQUE ID ##
         SCRAM,
         ## RESPONSE WEEK ##
         WEEK,
         ## DEMOGRAPHICS ##
         TBIRTH_YEAR,
         EGENDER, #phase 3.1
         EGENID_BIRTH, #phase 3.2+
         RHISPANIC,
         RRACE,
         EEDUC,
         ## DISABILITY ##
         SEEING,
         HEARING,
         REMEMBERING,
         MOBILITY,
         ## EMPLOYMENT ##
         ANYWORK,
         WKVOL,
         SETTING,
         ## VACCINATION ##
         RECVDVACC,
         ## LOCATION ##
         REGION,
         EST_ST,
         ## WEIGHTS ##
         PWEIGHT,
         HWEIGHT,
         `PWEIGHT1-PWEIGHT80`)
```

# 2.2 Clean data and create new variables - independent variables

```{r}

df_ind <- df_selected %>%
  mutate(
         #Age groups
         agegroup1 = 
                     as.character(
                     #For 2021 (Wk28-40) data
                     ifelse(28<=WEEK & WEEK<=40,
                       ifelse(2001<=TBIRTH_YEAR & TBIRTH_YEAR<=2003,1, #ages 18 to 20
                       ifelse(1972<=TBIRTH_YEAR & TBIRTH_YEAR<=2000,2, #ages 21 to 49
                       ifelse(                    TBIRTH_YEAR<=1971,3, #ages 50+
                     NA))),
                     #For 2022 (Wk41+) data
                       ifelse(2002<=TBIRTH_YEAR & TBIRTH_YEAR<=2004,1, #ages 18 to 20
                       ifelse(1973<=TBIRTH_YEAR & TBIRTH_YEAR<=2001,2, #ages 21 to 49
                       ifelse(                    TBIRTH_YEAR<=1972,3, #ages 50+
                     NA))))),
         agegroup_label = 
                     #For 2021 (Wk28-40) data
                     ifelse(28<=WEEK & WEEK<=40,
                       ifelse(2001<=TBIRTH_YEAR & TBIRTH_YEAR<=2003,'ages 18 to 20', #ages 18 to 20
                       ifelse(1972<=TBIRTH_YEAR & TBIRTH_YEAR<=2000,'ages 21 to 49', #ages 21 to 49
                       ifelse(                    TBIRTH_YEAR<=1971,'ages 50+', #ages 50+
                     NA))),
                     #For 2022 (Wk41+) data
                       ifelse(2002<=TBIRTH_YEAR & TBIRTH_YEAR<=2004,'ages 18 to 20', #ages 18 to 20
                       ifelse(1973<=TBIRTH_YEAR & TBIRTH_YEAR<=2001,'ages 21 to 49', #ages 21 to 49
                       ifelse(                    TBIRTH_YEAR<=1972,'ages 50+', #ages 50+
                     NA)))),
         
         #Gender
         female = 
                     as.character(
                     #For phase 3.1 (week 28-33) data
                     ifelse(28<=WEEK & WEEK<=33,
                       ifelse(EGENDER == 2,1,0
                      ),
                     #For phase 3.2-3.4 (week >=34) data
                       ifelse(EGENID_BIRTH  == 2,1,0)
                      )),
         #Hisp and Race
         hisprace = 
                     as.character(
                     ifelse(RHISPANIC == 2              ,1, #Hispanic
                     ifelse(RHISPANIC == 1 & RRACE == 1 ,2, #White, alone
                     ifelse(RHISPANIC == 1 & RRACE == 2 ,3, #Black, alone
                     ifelse(RHISPANIC == 1 & RRACE == 3 ,4, #Asian
                     ifelse(RHISPANIC == 1 & RRACE == 4 ,5, #Any other race alone, or in combination
                     NA)))))),
         hisp = 
                     as.character(
                     ifelse(RHISPANIC ==2                ,1,0)
                     ),
         black = 
                     as.character(
                     ifelse(RHISPANIC == 1 & RRACE == 2 ,1,0)
                     ),
         asian = 
                     as.character(
                     ifelse(RHISPANIC == 1 & RRACE == 3 ,1,0)
                     ),
         other = 
                     as.character(
                     ifelse(RHISPANIC == 1 & RRACE == 3 ,1,0)
                     ),
         #Education
         educ =
                     as.character(
                     ifelse(EEDUC %in% c(1,2) ,1, #Less than high school
                     ifelse(EEDUC == 1       ,2, #High school
                     ifelse(EEDUC %in% c(4,5) ,3, #Some college
                     ifelse(EEDUC %in% c(6,7) ,4, #College
                     NA))))),
         lesshs = 
                     as.character(
                     ifelse(EEDUC %in% c(1,2) ,1,0)
                     ),
         hs = 
                     as.character(
                     ifelse(EEDUC == 3 ,1,0)
                     ),
         somecol =
                     as.character(
                     ifelse(EEDUC %in% c(4,5) ,1,0)
                     ),
         bachplus =
                     as.character(
                     ifelse(EEDUC %in% c(6,7) ,1,0)
                     ),
         #Disability
         disability = 
                     as.character(
                     ifelse(SEEING %in% c(3,4) | 
                            HEARING %in% c(3,4) | 
                            REMEMBERING %in% c(3,4) | 
                            MOBILITY %in% c(3,4),1,0)
                     )
         )


#df_ind %>% filter(is.na(agebroup1))
#df_ind %>% filter(is.na(female))

df_ind %>% write.csv('PULSE.csv')
```